Exploring Real Data: A look at AirBnB

### 2017-06-07 ArticleDraft.Rmd
Mathmanda edited beyond the basics, intro and conclusion.  Added a formatted version of the word file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) # graphics
library(dplyr) # data wrangling
library(cowplot) #plot grids
library(knitr) # fancy tables
library(GGally) # scatterplot matrices
library(tree) # regression trees
library(rpart) #classification trees

@@

Data is seemingly everywhere these days, and is becoming more powerful as we learn how to use it to answer important questions. In this article we highlight one freely-available, rich and complex data set which you might use to try out asking and answering data-driven questions.

What follows are a collection of explorations that you can try out along with us.  After each exploration we’ll give some of our own thoughts, discussion, and visualization.  We encourage the reader to head over to insideairbnb.com, find some interesting data, and start exploring.

@@

## Moving Beyond the Basics 
In this section we pose more questions and leave the investigations open to you. If you would like to see how we answered these questions wth the Seattle data, you can visit our workfile at mathquest.carroll.edu/AirbnbExplorations/.  Several of these exercises suggest more advanced techniques, such as multiple regression, logistic regression, or machine learning techniques to answer.  A good idea for each of the following explorations would be to split your data into a training and a test set so that you can test your predictions on data that your model has not seen.

@@

###Exploration #6: 
What variables in this data set do you think would be the most useful in predicting the price of a particular rental?  Try `accommodates`, `beds`, `bathrooms`, `guests_included`, `review_scores_rating`, and various neighborhoods, and see if you can build a model (maybe a multiple regression, regression tree, or artificial neural network) that is *good* at predicting prices.

@@

```{r, echo = showcode, eval = F}
seattle1<-seattle %>% select(price, accommodates, beds, bathrooms,
                            guests_included,review_scores_rating, neighbourhood_cleansed)

#What are the most popular neighbourhoods? 
seattle1 %>% group_by(neighbourhood_cleansed) %>% tally() %>% arrange(desc(n))

nbd_list<-c('Broadway', 'Wallingford', 'Belltown')
seattle1<-seattle1 %>% filter(neighbourhood_cleansed %in% nbd_list)

#Let's look at how these variables relate to each other: 
ggpairs(seattle1, cardinality_threshold = 90)

#It looks like price is correlated with accomodates, beds, and baths ... but these are also correlated with each other! Does it seem to be related to neighbourhood?

seattle1 %>% group_by(neighbourhood_cleansed) %>% summarise(mean = mean(price), median = median(price))
#Belltown seems to cost more than the other two.  Let's create a dummy variable for Belltown.

seattle1<- seattle1 %>% mutate(BT=ifelse(neighbourhood_cleansed=='Belltown',1,0))

@@

m1<-lm(price~ accommodates+beds+bathrooms +guests_included+review_scores_rating+BT, data = seattle1 )
summary(m1)

#At first glance it appears that we can explain about 43% of the variation in price using these variables, and the ones that seem to contribute the most are 'accomodates', 'bathrooms, 'guests_included', and 'BT' (perhaps, in that order).  But... does it seem weird to you that 'beds' is not in this list? Can you explain why it's not? There are many ways to decide which variables to include (like stepwise regression, or contextual knowledge of the problem).  Let's naively just take the ones I mentioned above:

seattle2<-seattle1 %>% select(price, accommodates, bathrooms, guests_included, BT)
seattle2<-na.omit(seattle2)
m2<-lm(price~ accommodates+bathrooms +guests_included+BT, data = seattle2 )
summary(m2)
stres<-rstandard(m2)
#Interestingly, we can now explain something like 47% of the variation in price, suggesting some of the variables we dropped were maybe acting like red herrings!  If we choose to accept this model (which, by the way, we shouldn't -- at least not until we assess its validity!), a practical description of the model would be:

#An (empty) rental costs $0. 
# - For every person the rental accomodates, add $25.65 to the price
# - For every bathroom in the rental, add $35.69 to the price
# - For each guest included, add $7.27 o the rental price
# - If the rental is in BellTown, add $19 to the price

#Assess validity of model: Do the residuals appear to be normally distributed? 
p1<-ggplot(seattle2, aes(x = accommodates, y = rstandard(m2))) + geom_point(shape = 1)
p2<-ggplot(seattle2, aes(x = bathrooms, y = rstandard(m2))) + geom_point(shape = 1)
p3<-ggplot(seattle2, aes(x = guests_included, y = rstandard(m2))) + geom_point(shape = 1)
p4<-ggplot(seattle2, aes(x = lm.influence(m2)$hat, y = rstandard(m2))) + 
  geom_point(shape = 1) + 
  geom_vline(xintercept = 2*(5)/length(stres), color = 'cyan')+ 
  geom_hline(yintercept = -4, color = 'red')+ 
  geom_hline(yintercept = 4, color = 'red')
plot_grid(p1,p2,p3,p4)
ggplot(seattle2, aes(x = fitted(m2), y = price)) + geom_point(shape = 1)

#While there is no recognizable nonlinear shape in these residual plots, we do have a number of 'outlier + leverage' points that could be skewing things for us.  Also, the variation of residuals appears to be growing as accomodates/bathroom get larger. So, this model has some problems.

@@

#The following code produces a regression tree for this problem:
t<-tree(price~ accommodates+beds+bathrooms +guests_included+review_scores_rating+BT, data = seattle1 )
t
plot(t)
text(t) # *********** this line errors **************

#It appears that the most important variable in the data is 'accomodates'.  
#If you just want to split the data into two groups, you should look to see if the renatl accomodates 8 or more people.  If so, the average price would be $390, if not the average price would be $125.  
#For the rentals which accomodate less than 8 people, the next break occurs depending on if he rental accomodates four or more people (average price $163.10) or not (average price $102.20). #The next break occurs in the rentals that accommodate more than 8 people: rentals in Belltown (average price $539) or not (average price $342.70).

#We can keep following the 'breaks' in this regression tree until a good 'cut-off point'.  Finding the right cut-off point is an interesting problem which you should learn more about.  
# Further Questions: 
#1 - Can you interpret the meaning behind all of the breaks? 
#2 - Does the regression tree change your mind about which variables (in which cases) are practically significant? 
#3 - Which model do you think is better?

@@

seattle1<-seattle %>% select(room_type, price, accommodates, beds, bathrooms,
                            guests_included,review_scores_rating, neighbourhood_cleansed)
seattle1<- seattle1 %>% filter(room_type != 'Shared room')
seattle1<- seattle1 %>% mutate(Home = ifelse(room_type == 'Private room', 0, 1))

#Which neighborhoods have more Private rentals? 
seattle1 %>% group_by(neighbourhood_cleansed) %>% summarize(p = mean(Private))%>%
  arrange(p)

#Suppose we only use price to predict whether a rental is a private room (instead of an entire house)

m1<-glm(Home~ price, family = binomial, data = seattle1 )
c<-coefficients(m1)

f<-function(x){1/(1 + exp(-c[[1]] -c[[2]]*x))}

seattlebinned<-seattle1 %>% mutate(pricebin = cut(price, breaks = 20*0:100, labels = 20*0.5:99.5))
seattleb<-seattlebinned %>% group_by(pricebin) %>% summarise(p = mean(Home))
ggplot(seattleb, aes(x = currency_to_numeric(pricebin), y = p)) + geom_point()+
  stat_function(fun = f, color = 'blue') + xlab('Price')+xlim(0,500)
#The graphic deomonstrates that an increase of price is linked to an increased likelihood that the rental is an entire home.

@@

# Let's try adding more variables:

seattle2<-na.omit(seattle1)
m2<-glm(Home~ accommodates+beds+bathrooms +guests_included+review_scores_rating + price, family = binomial, data = seattle2 )
summary(m2)

m3<-glm(Home~ accommodates+bathrooms +review_scores_rating + price, family = binomial, data = seattle2 )
summary(m3)

seattle2<- seattle2 %>% mutate(Predict = ifelse(fitted(m3)>.5, 1, 0))
classError(seattle2$Home,seattle2$Predict )
#New error rate: 13%.  Not that much better?

@@

t<-rpart(Home~ accommodates+beds+bathrooms +guests_included+review_scores_rating + price, 
         data = seattle2 )
t
plot(t, compress = TRUE)
text(t, use.n = TRUE)

#Fix plot and add interpretation

@@

```{r, eval = F, echo = showcode}

seattle1<-seattle %>% select(room_type, price, accommodates, beds, bathrooms,
                            guests_included ,review_scores_rating,
                            neighbourhood_cleansed,latitude, longitude)
#Which nbds should we pick
seattle1 %>% group_by(neighbourhood_cleansed) %>% tally() %>% arrange(desc(n))
#Howabout: 
nbd_list<-c('Broadway', 'Belltown', 'Wallingford', 'University District', 'Minor')
seattle1<-seattle1 %>% filter(neighbourhood_cleansed %in% nbd_list)
ggplot(seattle1, aes(x = longitude, y = latitude, color = neighbourhood_cleansed)) + 
  geom_point()

seattle1 %>% tbl_df()

clusters<-seattle1 %>%
  mutate(Home = ifelse(room_type == 'Entire home/apt',1,0)) %>%
  select(-neighbourhood_cleansed,-room_type)

clusters<-na.omit(clusters)
km<-kmeans(cbind(clusters[,1:8]),5)
clusters<-clusters %>% mutate(Predict = km$cluster)

ggplot(clusters, aes(x = longitude, y = latitude, color = as.factor(Predict))) + 
  geom_point()

#Still need to play around with variables.  Or, just give up and use latitude and longitude only...

@@

##Conclusions
The Insideairbnb data set is robust enough that anyone with statistical training, no matter the sophistication, can ask meaningful and challenging questions.  In recent times the field of data science has become a successful and popular way to analyze data sets of this type. If you want to learn more about exploring and using data for analysis and predictions there are starter courses on the web (e.g., Coursera and DataCamp) and many universities and colleges around the world are implementing Data Science programs.  Most importantly, as you may have found from the last three exercises, the level of mathematical sophistication associated with data analysis can be quite high, so if you are a student reading this article and you find the ideas interesting then we recommend taking courses in statistics, computer science and computational mathematics.

There are many data sets like Insiderairbnb available for free on the internet.  For example, the University of California, Irnive, hosts a collection of machine learning data sets that are at least as rich and interesting (archive.ics.uci.edu/ml/).  Other sites to look at when seeking data sets include data.gov, kaggle.com, quandl.com, gapminder.org, flowingdata.com, and many others.  The world of data and data analysis is growing in importance socially and mathematicians and statisticians are uniquely positioned with their training to approach these questions with proper skills and mindsets.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) # graphics
library(dplyr) # data wrangling
library(cowplot) #plot grids
library(knitr) # fancy tables
library(GGally) # scatterplot matrices
library(tree) # regression trees
library(rpart) #classification trees
library(rpart.plot) #plots of classification trees
library(car) # marginal model plots

@@

According to 
Jonathan Rosenberg, adviser to Larry Page and former SVP of products at Google, 
"Data is the sword of the 21st century, those who weild it well, the samurai."
The world of data and data analysis is growing in importance, and those with the interest and appropriate skillset are uniquely positioned to solve interesting problems and answer vital questions.

In this article we highlight one freely-available, rich and complex data set which you might use to try out asking and answering data-driven questions. 
What follows are a collection of explorations that you can try out along with us.  After each exploration we’ll give some of our own thoughts, discussion, and visualization.  We encourage the reader to head over to insideairbnb.com, find some interesting data, and start exploring.

@@

## Moving Beyond the Basics

In this section we pose more questions and leave the investigations open to you. If you would like to see how we answered these questions wth the Seattle data, you can visit our workfile at mathquest.carroll.edu/AirbnbExplorations/.  Several of these exercises suggest more advanced techniques, such as multiple regression, logistic regression, or machine learning techniques to answer.  A good idea for each of the following explorations would be to split your data into a training and a test set so that you can test your predictions on data that your model has not seen.

@@

###Exploration #6:

What variables in this data set do you think would be the most useful in predicting the price of a particular rental?  For example, try `accommodates`, `beds`, `bathrooms`, `guests_included`, `review_scores_rating`, and dummy variables for various neighborhoods, and see if you can build a model (maybe a multiple regression, regression tree, or artificial neural network) that is good at predicting prices.

@@

```{r, echo = showcode, eval = F}
#First, let's keep only the variables we're interested in:
seattle1<-seattle %>% 
  select(price, accommodates, beds, bathrooms,
          guests_included,review_scores_rating, neighbourhood_cleansed)

# Let's look at how these variables relate to each other
# by creating a scatterplot matrix:
ggpairs(seattle1[,1:6])
#It looks like price is correlated with accomodates, beds, and baths ...
# but these are also correlated with each other!

# Does price seem to be related to neighbourhood? 
seattle1 %>% group_by(neighbourhood_cleansed) %>%
  summarise(mean = mean(price), median = median(price), n = n())%>%
  arrange(desc(n))

# Belltown seems to cost more than other neighborhoods.
# University District seems to cost less than other neighborhoods.
# Let's create dummy variables for Belltown and University District.

seattle1<- seattle1 %>% mutate(
  BT=ifelse(neighbourhood_cleansed=='Belltown',1,0),
  UD=ifelse(neighbourhood_cleansed=='University District',1,0))

@@

m1<-lm(price~ accommodates+beds+bathrooms+
         guests_included+review_scores_rating+BT+UD, data = seattle1 )
summary(m1)

#At first glance it appears that we can explain about 47% of the variation
#in price using these variables, and all variables seem to contribute to 
#price except 'beds'.

#Does it seem weird to you that 'beds' doesn't seem to contribute significantly? 
#Can you explain why it's not?

#There are many ways to decide which variables to include 
# (like stepwise regression, or contextual knowledge of the problem).  
# Let's naively just take the ones I mentioned above:

seattle2<-seattle1 %>% select(-beds)
seattle2<-na.omit(seattle2)
m2<-lm(price~ accommodates+bathrooms+guests_included+
         review_scores_rating+BT+UD, data = seattle2 )
summary(m2)
stres<-rstandard(m2)
#Interestingly, we can still explain about 47% of the variation in price, 
#suggesting that eliminating 'beds' doesn't cost us any predictive power.
# If we choose to accept this model 
# (which, by the way, we shouldn't -- at least not until we assess its validity!), 
# a practical description of the model would be:

# - For every person the rental accomodates, add $21.20 to the price
# - For every bathroom in the rental, add $35.15 to the price
# - For each guest included, add $6.52 o the rental price
# - For each point in the rating, add $.68 to the rental price
# - If the rental is in BellTown, add $30.47 to the price
# - If the rental is in University District, subtract $14.19 from the price

#What is the *practical* significance of the coefficients explained above?  
#For example, it seems like 'accommodates' and 'bathrooms' add a significant 
# amount of money to the price.  On the other hand, 'review_scores_rating' 
# *seems* statistically significant, adding 68 cents to the price of a rental 
# above $100 seems irrelevant (even if it is true).

#If you were going to open an Airbnb in Seattle, 
# which of the variables above would you consider, in order to maximize your profits?

#If you are interested in assessing the validity of our model, ask: 
# Do the residuals appear to be normally distributed? 
p1<-ggplot(seattle2, aes(x = accommodates, y = rstandard(m2))) + geom_point(shape = 1)+ xlab('# Accommodated') + ylab('St. Residuals')
p2<-ggplot(seattle2, aes(x = bathrooms, y = rstandard(m2))) + geom_point(shape = 1)+
   xlab('# Bathrooms') + ylab('St. Residuals')
p3<-ggplot(seattle2, aes(x = guests_included, y = rstandard(m2))) + 
  geom_point(shape = 1)+ xlab('# Guests Included') + ylab('St. Residuals')
p4<-ggplot(seattle2, aes(x = lm.influence(m2)$hat, y = rstandard(m2))) + 
  geom_point(shape = 1) + 
  geom_vline(xintercept = 2*(5)/length(stres), color = 'cyan')+ 
  geom_hline(yintercept = -4, color = 'red')+ 
  geom_hline(yintercept = 4, color = 'red')+ xlab('Influence Score') + ylab('St. Residuals')
plot_grid(p1,p2,p3,p4)
ggplot(seattle2, aes(x = fitted(m2), y = rstandard(m2))) + geom_point(shape = 1)+
  xlab('Fitted Values') + ylab('St. Residuals')

#While there is no recognizable nonlinear shape in these residual plots, 
# we do have a number of 'outlier + leverage' points that are problematic.  
# Also, the variation of residuals appears to be growing as things get larger. 
# So, this model is not without its problems.

@@

#Produces a regression tree:
t<-tree(price~ accommodates+beds+bathrooms +guests_included+review_scores_rating+BT+UD, data = seattle1 )
t
plot(t)
text(t)

#It appears that the most important variable in the data is 'accomodates'.  
#If you just want to split the data into two groups, you should look to
# see if the rental accomodates 5 or more people.  
#     If so, the average price would be $223, 
#     If not the average price would be $102.

#For the rentals which accomodate less than 5 people, 
# the next break occurs depending on 
#    if he rental accomodates three or four people (average price $128.10) 
#    or not (average price $85.51).

#The next break occurs in the rentals that accommodate 5 or more people: 
#     rentals with two or fewer bathrooms (average price $188) 
#     or more than two (average price $326).

#We can keep following the 'breaks' in this regression tree until a good 'cut-off point'.  Finding the right cut-off point is an interesting problem which you should learn more about.  
# Further Questions: 
#1 - Can you interpret the meaning behind all of the breaks? 
#2 - How many 'breaks' in the tree would you consider necessary, 
#     before you might be 'overfitting' the data? Why?
#3 - Does this regression tree change your mind about which variables (in which cases) are practically significant? 
#4 - Which model we've produced so far do you think is better for predicting price?

@@

# First, we consider only the variables we care about: 
seattle1<-seattle %>% select(room_type, price, accommodates, beds, bathrooms,
                            guests_included,review_scores_rating,
                            neighbourhood_cleansed)

#For simplicity, let's eliminate the Shared rooms, as they are irrelevant here
seattle1<- seattle1 %>% filter(room_type != 'Shared room')
seattle1<- seattle1 %>% mutate(Home = ifelse(room_type == 'Private room', 0, 1))

#Suppose we only use price to predict whether a rental is a private room (instead of an entire house)
m1<-glm(Home~ price, family = binomial, data = seattle1 )
c<-coefficients(m1)
#We create the logistic function that has been estimated: 
f<-function(x){1/(1 + exp(-c[[1]] -c[[2]]*x))}

#In order to create a visualization of what we've just done, we're going to bin our data into buckets of observations that have similar prices, and compare their proportion of 'entire home' rentals':
seattlebinned<-seattle1 %>% 
  mutate(pricebin = cut(price, breaks = 20*0:100, labels = 20*0.5:99.5))
seattleb<-seattlebinned %>% group_by(pricebin) %>% summarise(p = mean(Home))
ggplot(seattleb, aes(x = currency_to_numeric(pricebin), y = p)) + geom_point()+
  stat_function(fun = f, color = 'blue') + xlab('Price')+xlim(0,500)
#The graphic deomonstrates that an increase of price is linked to an increased likelihood that the rental is an entire home.

@@

# Let's try adding more variables: 
seattle2<-na.omit(seattle1)
m2<-glm(Home~ accommodates+beds+bathrooms +guests_included+review_scores_rating + price, family = binomial, data = seattle2 )
summary(m2)

#Somewhat arbitrarily, I've decided to keep accomodates, bathrooms, and price: 
m3<-glm(Home~ accommodates+bathrooms  + price, family = binomial, data = seattle2 )
summary(m3)

#How effectively does this predict 'entire home's? 
seattle2<- seattle2 %>% mutate(Predict = ifelse(fitted(m3)>.5, 1, 0))
classError(seattle2$Home,seattle2$Predict )
#New error rate: 13.8%.  Not that much better?

@@

#Now, let's produce a classification tree for this question: 
t<-rpart(Home~ accommodates+beds+bathrooms +guests_included+
           review_scores_rating + price, data = seattle2 )
t
rpart.plot(t, compress = TRUE)

#1 - Can you interpret the meaning behind all of the breaks in this tree? 
#2 - How many 'breaks' in the tree would you consider necessary, 
#     before you might be 'overfitting' the data? Why?
#3 - Does this classification tree change your mind about which variables 
#    (and in which cases) are practically significant? 
#4 - Which model we've produced so far do you think is better for 
#    predicting Homes vs. private rooms?

@@

```{r, eval = F, echo = showcode}
#As usual, let's start by isolating the variables we care about: 
seattle1<-seattle %>% select(room_type, price, accommodates, beds, bathrooms,
                            guests_included ,review_scores_rating,
                            neighbourhood_cleansed,latitude, longitude)

#We've decided to use the following 5 neighbourhoods:  
nbd_list<-c('Broadway', 'Belltown', 'University District', 'Minor','Fremont')
#Let's eliminate all rentals that aren't in these neighborhoods: 
seattle1<-seattle1 %>% filter(neighbourhood_cleansed %in% nbd_list)
ggplot(seattle1, aes(x = longitude, y = latitude, color = neighbourhood_cleansed)) + 
  geom_point()

#Now we create a new data set with a dummy variable for 'entire home' and 
# elminating non-numeric variables
clusters<-seattle1 %>%
  mutate(Home = ifelse(room_type == 'Entire home/apt',1,0)) %>%
  select(-neighbourhood_cleansed,-room_type)%>%
  na.omit()

#K-means is a way to determining clusters in data by trying to minimize
# 'within-cluster-sum-of-squares'
#What if we try to detect just 2 clusters here 
km<-kmeans(cbind(clusters[,1],clusters[,7:8]),3)
clusters<-clusters %>% mutate(Predict = km$cluster)
ggplot(clusters, aes(x = longitude, y = latitude, color = as.factor(Predict))) + 
  geom_point()

#As you play around with the code above you may notice a few of the following things: # - Each time we xecute the code above we get potentially different clusters, as this method depends on randomly chosen starting locations
# - Some number of clusters work well, if we guess at too many or too few, we get bad results
# - Latitude and Longitude get almost ignored when any other variable is tossed into the model.  Is this because the variability in lat/long is so small compared to other variables? If so, can you think of a way to compat that?

#Let's try using only lat/long, and trying to decide what the optimal number of clusters is: 
wcc<-NULL
for(i in 2:10){
  wcc<-c(wcc,sum(kmeans(clusters[,7:8],i)$withinss))	
}
plot(seq(2,10),wcc,type="b",pch=1,xlab="# of Clusters",
     ylab="Within Sum--Squares")
#In the 'elbow plot' above, we look for places where the slope changes drastically

#Suppose you choose 4 clusters: 
km<-kmeans(clusters[,7:8],4)
clusters<-clusters %>% mutate(Predict = km$cluster)
ggplot(clusters, aes(x = longitude, y = latitude, color = as.factor(Predict))) + 
  geom_point()

#How could you improve our clustering model? 
#Can you think of a way to not rely on latitude and longitude?

@@

##Conclusions

<!--The Insideairbnb data set is robust enough that anyone with statistical training, no matter the sophistication, can ask meaningful and challenging questions.-->  Recently, the field of data science has become a successful and popular way to analyze data like the one presented in this article. If you want to learn more about this field, there are starter courses on the web (e.g., Coursera and DataCamp) and many universities and colleges around the world are implementing data science programs.  As you may have discovered in the last three exercises, the level of mathematical sophistication associated with data analysis can be quite high, so if you find the data-driven problems interesting, then we recommend taking more courses in statistics, computer science and computational mathematics! For, as  Jennifer Pahlka, Founder and Executive Director for Code for America has said: "[o]ur ability to do great things with data will make a real difference in every aspect of our lives."

Many data sets like the one we've considered here are available for free online.  
For example, some of our favorite places to find interesting (and real) data are

  * data.gov
  * kaggle.com
  * archive.ics.uci.edu/ml
  * Quandl.com
  * gapminder.org
  * flowingdata.com

<!--For example, the University of California, Irnive, hosts a collection of machine learning data sets that are at least as rich and interesting (archive.ics.uci.edu/ml/).  Other sites to look at when seeking data sets include data.gov, kaggle.com, quandl.com, gapminder.org, flowingdata.com, and many others.  -->
